---
title: 'Predictive Modeling Process'
author: "Aoyi Shan, Yukun He"
date: "11/1/2016"
output: ioslides_presentation
---

# Introduction

In this project, we explore the difference between various muitiple regression methods and conclude on what approach should be used in order to fit our data better. In our analysis, we compare the fit of models generated by 5 different regression methods, regression model obtained by Ordinary Least Squares, Ridge regression, Lasso regression, Principal Components regression and Partial Least Squares regression. To evaluate how well each model fits the data, we perform 10-fold cross validation in the model construction process and select the model with minimum cross-validation errors in terms of the running parameter.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

